\pkg{sdmTMB}: An \proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields

### 2023-09-15 doc/paper-jss/sdmTMB-paper.Rnw
Edits

\sloppy We then create a mesh object that contains triangulation and projection
matrices needed to apply the SPDE approach. The argument \texttt{cutoff} defines
the minimum allowed distance between mesh vertices in the units of \texttt{X}
and \texttt{Y}. \texttt{make\_mesh()} is a wrapper function for
\texttt{INLA::inla.mesh.create()} or \texttt{fmesher::fm\_rcdt\_2d\_inla()} and
\texttt{cutoff} is passed through to these functions. Alternatively, we could
have created any mesh via \proglang{R}-\pkg{INLA} or \pkg{fmesher} and supplied
it to the \texttt{mesh} argument in \texttt{make\_mesh()}. We recommend
constructing such a mesh for applied analyses to take advantage of
additional functionality such as coarser mesh triangles in a border around the
data and control over maximum triangle length and minimum triangle angles. We
can inspect mesh object with the associated plotting method. Our mesh has
238 (\texttt{mesh\_pcod\$mesh\$n}) vertices.

<<dog-binomial-mesh, results='hide', message=FALSE, warning=FALSE, echo=TRUE, fig.cap="SPDE mesh (lines) combined with the trawl survey observations (points). The locations where lines intersect are referred to as ``vertices''. Finer meshes will be slower to fit but generally increase the accuracy of the SPDE approximation. This is a simple mesh created with the convenience function \\code{make\\_mesh()}. A greater degree of control over the mesh construction can be achieved by using the \\pkg{fmesher} or \\proglang{R}-\\pkg{INLA} directly.", fig.width=5>>=
mesh_pcod <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 8)
plot(mesh_pcod)
@

@@

We can also fit a version that omits the spatial random field by setting
\texttt{spatial = "off"}. We will use the \texttt{update()} S3 method to refit
the model updating any specified arguments:

@@

Here, this does not flag any issues. \texttt{sanity()} is checking that the
\texttt{nlminb()} optimizer reported successful convergence, that the Hessian
matrix is positive definite, that no extreme or small eigenvalues are detected,
that no absolute log likelihood gradients with respect to fixed defects are
$\le$ 0.001, that all fixed effects have reported standard errors that do not
look unreasonably large ($<$ 100 by default), that random field marginal
standard deviations are not unexpectedly small or large ($>$
0.01 and $<$ 100), and that the random field Matérn range parameter does not
  look unreasonably large ($<$ 1.5 times the largest edge of a bounding box
  around the observations).

@@

We can also see that the AIC of the model with spatial random fields is
considerably lower:

@@

We can do additional model comparison with cross validation via
\texttt{sdmTMB\_cv()}. Here we will do 10-fold cross validation with the folds
constructed randomly. Using the \texttt{fold\_ids} argument, we could supply our
own folds and conduct spatially blocked cross validation. If we set a parallel
plan with the \pkg{future} (REF TODO) package, our folds will be fit in
parallel.

@@

\subsection{Visualizing model predictions}

>>>>

\sloppy We then create a mesh object that contains triangulation and projection
matrices needed to apply the SPDE approach. The argument \texttt{cutoff} defines
the minimum allowed distance between mesh vertices in the units of \texttt{X}
and \texttt{Y}. \texttt{make\_mesh()} is a wrapper function for
\texttt{INLA::inla.mesh.create()} or \texttt{fmesher::fm\_rcdt\_2d\_inla()} and
\texttt{cutoff} is passed through to these functions. Alternatively, we could
have created any mesh via \proglang{R}-\pkg{INLA} or \pkg{fmesher} and supplied
it to the \texttt{mesh} argument in \texttt{make\_mesh()}. We recommend
constructing such a mesh for applied analyses to take advantage of additional
functionality such as coarser mesh triangles in a border around the data and
control over maximum triangle length and minimum triangle angles. We can inspect
mesh object with the associated plotting method
(Figure~\ref{fig:dog-binomial-mesh}). Our mesh has 238
(\texttt{mesh\_pcod\$mesh\$n}) vertices.

<<dog-binomial-mesh, results='hide', message=FALSE, warning=FALSE, echo=TRUE, fig.cap="SPDE mesh (lines) combined with the trawl survey observations (points). The locations where lines intersect are referred to as ``vertices''. Finer meshes will be slower to fit but generally increase the accuracy of the SPDE approximation. This is a simple mesh created with the convenience function \\code{make\\_mesh()}. A greater degree of control over the mesh construction can be achieved by using \\pkg{fmesher} or \\proglang{R}-\\pkg{INLA} directly.", fig.width=5>>=
mesh_pcod <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 8)
plot(mesh_pcod)
@

@@

We can also fit a version that omits the spatial random field by setting
\texttt{spatial = "off"}. We will use the \texttt{update()} S3 method to refit
the model while updating any specified arguments:

@@

This does not flag any issues. \texttt{sanity()} is checking that the
\texttt{nlminb()} optimizer reported successful convergence, that the Hessian
matrix is positive definite, that no extreme or small eigenvalues are detected,
that no absolute log likelihood gradients with respect to fixed defects are
$\le$ 0.001, that all fixed effects have reported standard errors that do not
look unreasonably large ($<$ 100 by default), that random field marginal
standard deviations are not unexpectedly small or large ($>$
0.01 and $<$ 100), and that the random field Matérn range parameter does not
  look unreasonably large ($<$ 1.5 times the largest edge of a bounding box
  around the observations).

@@

We can also see that the AIC of the model with spatial random fields is
lower:

@@

We can do additional model comparison with cross validation via
\texttt{sdmTMB\_cv()}. Here we will do 10-fold cross validation with the folds
constructed randomly. Using the \texttt{fold\_ids} argument, we could supply our
own folds and conduct spatially blocked cross validation. If we set a parallel
plan with the \pkg{future} package \citep{bengtsson2021future}, our folds will
be fit in parallel.

@@

\subsection{Making predictions on new data}
